# Reproducible Research: Peer Assessment 1
Research reproduced by: Zhengming Song at `r date()`

```{r setoptions,echo=FALSE}
opts_chunk$set(fig.height=6,fig.width=11.3)
```
## Loading and preprocessing the data
```{r,cache=TRUE}
rawdata = read.table("activity.csv",sep=",",header = T)
data = rawdata[!is.na(rawdata$steps),]
dates = levels(data$date)
cleandata = data.frame(date = as.Date(dates),totsteps = rep(0,length(dates)))
for(i in 1:length(dates)){
  cleandata$totsteps[i]=sum(data$steps[data$date == dates[i]])
}
head(cleandata)
```

## What is mean total number of steps taken per day?
```{r}
hist(cleandata$totsteps, xlab="steps")
avg = mean(cleandata$totsteps)
med = median(cleandata$totsteps)
```
Mean of total number of steps taken per day is `r avg`  
Median of total number of steps taken per day is `r med`

## What is the average daily activity pattern?
```{r}
intervals = levels(factor(data$interval))
avgstep = rep(0,length(intervals))
for(i in 1:length(intervals)){
  avgstep[i]=mean(data$steps[data$interval == intervals[i]])
}

plot(intervals, avgstep,type="l")
```
Determine which interval contains the maximum number of steps on average across all the days
```{r}
ind = intervals[which(avgstep == max(avgstep))]
```
It turns out interval `r ind` contains maximum number of steps on average across all the days

## Imputing missing values
```{r}
num_na = sum(is.na(rawdata$steps))
```
* There are `r num_na` rows with NAs
* Let's fill the missing values by the average value of that 5-minute interval
* Let's call the new dataset as newdata
```{r}
newdata = rawdata
ind = which(is.na(newdata),arr.ind=TRUE)
for (i in 1:dim(ind)[1]){
  newdata$steps[ind[i,1]] = avgstep[intervals == newdata$interval[ind[i,1]]]
}
head(newdata)
```

```{r}
dates = levels(newdata$date)
cleandata = data.frame(date = as.Date(dates),totsteps = rep(0,length(dates)))
for(i in 1:length(dates)){
  cleandata$totsteps[i]=sum(newdata$steps[newdata$date == dates[i]])
}
hist(cleandata$totsteps, xlab="steps")
avg = mean(cleandata$totsteps)
med = median(cleandata$totsteps)
```
* The mean value is `r avg`, and the median is `r med` now. we can observe that the histogram and values are different as those from first part. Filling the missing data fix the estimation bias (the mean changed)

## Are there differences in activity patterns between weekdays and weekends?
```{r}
# Sepreate newdata to two sub dataset according to weekday or weekend
dates = as.Date(newdata$date)
weekendsind = which(weekdays(dates) == "Sunday" | weekdays(dates) == "Saturday", arr.ind=TRUE)
weekdaysind = which(weekdays(dates) == "Monday" | weekdays(dates) == "Tuesday" | weekdays(dates) == "Wednesday" | weekdays(dates) == "Thursday" | weekdays(dates) == "Friday", arr.ind=TRUE)
weekdaydata = newdata[weekdaysind,]
weekenddata = newdata[weekendsind,]

# construct new dataset, weekdata, to store the data for plotting
intervals = levels(factor(weekenddata$interval))
avgstep = rep(0,length(intervals))
for(i in 1:length(intervals)){
  avgstep[i]=mean(weekenddata$steps[weekenddata$interval == intervals[i]])
}

weekdata  = data.frame(interval= as.numeric(intervals), steps = avgstep)

intervals = levels(factor(weekdaydata$interval))
avgstep = rep(0,length(intervals))
for(i in 1:length(intervals)){
  avgstep[i]=mean(weekdaydata$steps[weekdaydata$interval == intervals[i]])
}

weekdata = rbind(data.frame(interval = as.numeric(intervals),steps = avgstep),weekdata)

# plot for comparison
library(lattice) 
attach(weekdata)
weekday.f = gl(2,length(interval)/2,labels = c("weekday","weekend"))
xyplot(steps~interval|weekday.f, auto.key=TRUE,layout = c(1,2),type='l', ylab = "Number of steps")
```
* The new factor variable is weekday.f
* From the panel plot above, we can observe that the behavior is different from weekend to weekday.

